---
title: "AirBnB Market Trend"
format: html
---
As a consultant working for a real estate start-up, you have collected Airbnb listing data from various sources to investigate the short-term rental market in New York. You'll analyze this data to provide insights on private rooms to the real estate company.
```{r}
# Load packages
library(tidyverse)
library(readxl) #  For reading Excel files (.xlsx or .xls)
```

```{r}
# Load datasets
# Load CSV file for prices
airbnb_price <- read_csv('data/airbnb_price.csv', show_col_types = FALSE)

# Load Excel file for room types
airbnb_room_type <- readxl::read_xlsx('data/airbnb_room_type.xlsx')

# Load TSV file for review dates
airbnb_last_review <- read_tsv(
  'data/airbnb_last_review.tsv',
  show_col_types = FALSE
)
```

What are the dates of the earliest and most recent reviews?
```{r}
# Check data type
glimpse(airbnb_last_review)

# Convert the 'last_review' character column into date object
airbnb_last_review = airbnb_last_review %>%
  mutate(last_review_date = parse_date_time(last_review, orders = "B d Y"))

first_reviewed <- min(airbnb_last_review$last_review_date, na.rm = TRUE)
last_reviewed <- max(airbnb_last_review$last_review_date, na.rm = TRUE)
```

How many of the listings are private rooms?
```{r}
head(airbnb_room_type)
unique(airbnb_room_type$room_type)

# Normalize and List Unique Room Types
airbnb_room_type = airbnb_room_type %>%
  mutate(room_type_clean = str_to_title(str_trim(room_type)))

airbnb_room_type %>%
  distinct(room_type_clean)

# Total Private room listings
nb_private_rooms = airbnb_room_type %>%
  filter(room_type_clean == "Private Room") %>%
  summarise(nb_private_rooms = n()) |>
  pull()
```

What is the average listing price for all rooms (rounded to the nearest penny)?
```{r}
head(airbnb_price)
glimpse(airbnb_price)

airbnb_price_clean <- airbnb_price %>%
  mutate(
    price_in_dollars = str_extract(price, "\\d+") %>%
      as.numeric()
  )

# Average listing price for all rooms
avg_price = airbnb_price_clean %>%
  summarise(avg_price = round(mean(price_in_dollars), 2)) |>
  pull()
```

```{r}
## Combine these four values into one tibble called review_dates with four columns: first_reviewed, last_reviewed, nb_private_rooms, and avg_price. The tibble should only contain one row of values.

review_dates <- tibble(
  first_reviewed,
  last_reviewed,
  nb_private_rooms,
  avg_price
)

review_dates
```
