---
title: "Hypothesis testing"
format: html
---

Perform an appropriate hypothesis test to determine the p-value, and decide whether to reject the null hypothesis that the mean number of goals scored in women's international soccer matches is the same as men's. Use a 10% significance level.

Store the p-value and the test result in a data frame called result_df, withp_val as the p-value and result as either the string "fail to reject" or "reject", based on the test outcome.

# Hypotheses:
# H0: The distribution of goals is the same for women's and men's matches.
# HA: Goals in women's matches are stochastically greater than men's.
# Significance Level (alpha): 0.10

```{r}
# Import libraries
library(tidyverse)
library(patchwork)  # For combining plots

```

```{r}
# Load Datasets
men_results = read_csv("Data/men_results.csv")
women_results = read_csv("Data/women_results.csv")
```

### Filter datasets
```{r}
# Filter for FIFA World Cup match data for both men’s and women’s tournaments to include only matches after 2002-01-01.

glimpse(men_results) # Check data types
glimpse(women_results)

men = men_results |>
  filter(tournament == "FIFA World Cup" & date > "2002-01-01") |>
  mutate(total_goals = home_score + away_score)  # # Calculate total goals per match

women = women_results |>
  filter(tournament == "FIFA World Cup" & date > "2002-01-01") |>
  mutate(total_goals = home_score + away_score)

```

### Check for distribution
```{r}
# Plotting histograms to check the distribution (checking for normality).
men_plot <- ggplot(men, aes(x = total_goals)) +
  geom_histogram(fill = "tan3", bins = 15, color = "black") + 
  ggtitle("Goals Scored (Men's World Cup Matches)") +
  xlab("Total Goals Scored") +
  ylab("Frequency")

women_plot <- ggplot(women, aes(x = total_goals)) +
  geom_histogram(fill = "steelblue2", bins = 15, color = "black") + 
  ggtitle("Goals Scored (Women's World Cup Matches)") +
  xlab("Total Goals Scored") +
  ylab("Frequency")

combined_plot <- men_plot + women_plot
combined_plot

# Goal count data is right-skewed, suggesting a non-parametric test 
```

###  Run Wilcoxon Rank Sum Test
```{r}
# Non-parametric Wilcoxon Rank Sum Test (Mann-Whitney U) is used, as the distribution is non-normal (as confirmed by the histograms), 

# The 'wilcox.test' tests the median/distribution, which is an appropriate non-parametric 
# equivalent for the mean comparison when the data is not normal.
# alternative = "greater" tests if goals in women's (x) are greater than men's (y).

test_result <- wilcox.test( 
  x = women$total_goals,
  y = men$total_goals,
  alternative = "greater"
)

# Print the test output 
print(test_result)
```

```{r}
# Store result
# Set significance level
alpha = 0.10

# Extract p-value 
# Round to 4 decimal places
p_val <- round(test_result$p.value, 4)

# Determine the decision based on the p-value and the alpha (0.10)
result <- ifelse(p_val <= alpha, "reject", "fail to reject")

# Create result data frame
result_df <- data.frame(p_val = p_val, result = result)
result_df
```

