---
title: "Hypothesis testing"
format: html
---

Perform an appropriate hypothesis test to determine the p-value, and decide whether to reject the null hypothesis that the mean number of goals scored in women's international soccer matches is the same as men's. Use a 10% significance level.

Store the p-value and the test result in a data frame called result_df, withp_val as the p-value and result as either the string "fail to reject" or "reject", based on the test outcome.

```{r}
# Import libraries
library(tidyverse)
library(patchwork)

```

```{r}
# Load Datasets
men_results = read_csv("Data/men_results.csv")
women_results = read_csv("Data/women_results.csv")
```

### Filter datasets
```{r}
# Filter for FIFA World Cup match data for both men’s and women’s tournaments to include only matches after 2002-01-01.

glimpse(men_results) # Check data type
glimpse(women_results)

men = men_results |>
  filter(tournament == "FIFA World Cup" & date > "2002-01-01") |>
  mutate(total_goals = home_score + away_score)

women = women_results |>
  filter(tournament == "FIFA World Cup" & date > "2002-01-01") |>
  mutate(total_goals = home_score + away_score)

```

### Check for normality
```{r}
men_plot <- ggplot(men, aes(x = total_goals)) +
  geom_histogram(fill = "tan3", bins = 30) +
  ggtitle("Goals Scored (Men's)") +
  xlab("Goals Scored") +
  ylab("Frequency")

women_plot <- ggplot(women, aes(x = total_goals)) +
  geom_histogram(fill = "steelblue2", bins = 30) +
  ggtitle("Goals Scored (Women's)") +
  xlab("Goals Scored") +
  ylab("Frequency")

combined_plot <- men_plot + women_plot
combined_plot
```

###  Run  t-test 
```{r}

wilcox.test(
  x = women$total_goals,
  y = men$total_goals,
  alternative = "greater"
)
```

```{r}
# Run the test and store result
test_result <- wilcox.test(
  x = women$total_goals,
  y = men$total_goals,
  alternative = "greater"
)

# Extract p-value and decision
p_val <- round(test_result$p.value, 4)
result <- ifelse(p_val <= 0.10, "reject", "fail to reject")

# Create result data frame
result_df <- data.frame(p_val = p_val, result = result)
result_df
```

